Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-11 ms.tex
fixed discussion

The unresolved core X-ray luminosities of galactic nuclei provide a
powerful diagnostic of the SMBH accretion rates and how they vary
across different galaxy properties and SMBH masses (e.g.,
\citealt{Ho08} and references therein).  Figure~\ref{fig:miller} shows
our predictions for $\langle L_{X} \rangle =\epsilon_X \epsilon_{\rm
  bol} \dot{M} c^2$ in Eddington units as a function of black hole
mass, where $\epsilon_{\rm bol} = 0.1$ is the approximate bolometric
correction at X-ray energies appropriate for low luminosity AGN
(\citealt{Ho08}).  We have averaged the accretion rate used in
calculating $\langle L_X \rangle$ between the values for galaxies with
core ($\Gamma = 0.1$) and cusp ($\Gamma = 0.8$) stellar density
profiles using the average value of the inner density slope $\Gamma$
as a function of black hole mass as fit from the
\citet{LauerFaber+:2007a} galaxy sample, an exercise that results in
the relationship $\Gamma = 0.3 \Mbheight^{-0.24}$.

The left panel of Figure \ref{fig:miller} is calculated assuming a
constant low value for the radiative efficiency $\epsilon_X = 10^{-4}$
typical of those estimated for low luminosity AGN (e.g.,
\citealt{Ho:2009a}), while the right panel luminosities are calculated
instead assuming an efficiency of $\epsilon_X= 10 \dot{M}/\dot{M}_{\rm
  edd}$ that depends on the Eddington ratio as predicted by some
models for radiatively inefficient accretion flows
(\citealt{Narayan&Yi95}; \citealt{Narayan+98}).  Shown for comparison
are the X-ray measurements (green stars) and upper limits (gray
triangles) of early-type galaxies compiled by \citet{Miller+15}
(cf.~\citealt{Gallo+10}).  A green line shows the best power-law fit
to the X-ray luminosity from \citet{Miller+15}, given by $\langle
L_X/L_{\rm edd}\rangle \sim \Mbh^\alpha$ with $\alpha = -0.2$ (see also
\citep{Zhang+09, Pellegrini10, Gallo+10} for similar conclusions).
Yellow, teal, and red lines show the maximum accretion rates,
respectively, for thermally stable accretion (eq.~[\ref{eq:Mdotmax}]),
as set by SN Ia blow-out (eq.~[\ref{eq:eddr_Ia}]), and as allowed by
Compton heating feedback (eq.~[\ref{eq:MdotC}]).

@@

If heating in the nuclei of elliptical galaxies is similar to that
produced by the average star formation history of galaxies with
similar mass, then to first order we predict that $\langle
L_{X}/L_{\rm edd}\rangle$ should increase with black hole mass, though
only weakly for $M_{\bullet} \lesssim 3\times 10^{8}M_{\odot}$.  This
result is in tension with the observed down-sizing trend that the
average Eddington ratio is observed to decrease (albeit also weakly)
with $M_{\bullet}$.  However, in making such comparisons it is
important to keep in mind the enormous uncertainty in calculating the
luminosity of the accretion flow close to the black hole in a single
waveband based on the accretion rate on larger scales.  Black hole
mass-dependent efficiency factors could enter that we are not taking
into account.  For instance, if disk outflows carry away a substantial
fraction of the infalling mass before it reaches sufficiently small
radii to produce the observed X-rays
(e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}), then the
angular momentum of the infalling gas (and hence that of stars in the
galactic nucleus) might also influence the X-ray luminosity, in a way
that could depend systematically on $M_{\bullet}$.

A more readily addressable test is whether a steady-state accretion
picture is capable of explaining the large scatter of typically 2-3
orders of magnitude in the observed distribution of $L_{X}/L_{\rm
  edd}$ at fixed $M_{\bullet}$.  Scatter is expected to result from at
least two effects: (1) the strong sensitivity of the accretion rate to
the stellar wind velocity, $\dot{M}/\dot{M}_{\rm edd} \propto
v_{w}^{-3.8(-2.4)}$ for core(cusp) galaxies, respectively
(Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]) in
combination with the significant dependence of $v_w$ on stochastic
intermittency in the star formation history (~\ref{fig:NickPlot2}) and
(2) angular momentum of the infalling gas and consequentiallly the
size of small scale outflows may vary from galaxy to galaxy at fixed
$\Mbh$. The combination of (1) and (2) is capable of explaining much
of the observed sample, especially considering that scatter in $L_{\rm
  X}$ due to variations in $\dot{M}$ is exacerbated by the expectation
that $\epsilon_{X} \propto \dot{M}$ for radiatively inefficient
flows. At a wind heating parameter there is an order of magnitude
difference between $\dot{M}$ for core and cusp galaxies at fixed $v_w$
and $M_{\bullet}$ (Fig.~\ref{fig:bh_growth}) which could also lead to
scatter of $\dot{M}$ due to the variation in the Nuker stellar density
profile $\Gamma$ at fixed $M_{\bullet}$. However, \citet{Pellegrini10}
find that the scatter in $L_X$ is similar if one just looks at
core/cusp galaxies separately. There could also be scatter introduced
in using scaling relationships ($\Mbh-\sigma$) and ($\Mbh-L_{\rm
  bulge}$) to measure $\Mbh$ as \citet{Miller+15} do.

@@

Many of the upper
limits on $L_{\rm X}$ at low $M_{\bullet}$ may in fact correspond to a
separate ``thermally stable" population.  This possible dichotomy
calls into question the use of simple extrapolations of power-law fits
to the $L_{\rm X}(M_{\bullet}$) relationship to low $M_{\bullet}$ to
constrain the occupation fraction of black holes in low mass galaxies
(\citealt{Miller+15}).

>>>>

The (unresolved) core X-ray luminosities of galactic nuclei provide a
powerful diagnostic of the SMBH accretion rates and how they vary with SMBH mass and other galaxy properties (e.g.,
\citealt{Ho08} and references therein).  Figure~\ref{fig:miller} shows
our predictions for $\langle L_{X} \rangle =\epsilon_X \dot{M} c^2$ as a function of black hole
mass, where $\epsilon_{X} = \epsilon\epsilon_{\rm bol}$, $\epsilon$ is the total radiative efficiency, and $\epsilon_{\rm bol} \sim 0.1$ is the bolometric
correction into the measured X-ray band for low luminosity AGN (\citealt{Ho08}).  The value of $\langle L_X \rangle$ is calculated using the average accretion for galaxies with
core ($\Gamma = 0.1$) and cusp ($\Gamma = 0.8$) stellar density
profiles, using the relationship $\Gamma = 0.3 \Mbheight^{-0.24}$ fit to the \citet{LauerFaber+:2007a} galaxy sample.

The left panel of Figure \ref{fig:miller} is calculated assuming a
constant low value for the radiative efficiency of $\epsilon_X = 10^{-4}$,
typical of those estimated for low luminosity AGN (e.g.,
\citealt{Ho:2009a}), while the right panel luminosities are calculated
instead assuming an efficiency of $\epsilon_X= 10 \epsilon_{\rm bol}\dot{M}/\dot{M}_{\rm
  edd}$ that depends on the Eddington ratio as predicted by some
models for radiatively inefficient accretion flows
(\citealt{Narayan&Yi95}; \citealt{Narayan+98}).  Shown for comparison
are the X-ray measurements (green stars) and upper limits (gray
triangles) from the sample of early-type galaxies compiled by \citet{Miller+15}
(cf.~\citealt{Gallo+10}).  A green line shows the best power-law fit
to the X-ray luminosity from \citet{Miller+15}, given by $\langle
L_X/L_{\rm edd}\rangle \sim \Mbh^\alpha$ with $\alpha = -0.2$ (see also
\citep{Zhang+09, Pellegrini10, Gallo+10}).  Also shown are the maximum accretion rates,
respectively, for thermally stable accretion (eq.~[\ref{eq:Mdotmax}]; yellow),
as set by SN Ia blow-out (eq.~[\ref{eq:eddr_Ia}]; teal), and as allowed by
Compton heating feedback (eq.~[\ref{eq:MdotC}]; red).

@@

If the nuclei of elliptical galaxies are heated in a way similar to that due to the average star formation history of galaxies with
similar mass, then to first order we predict that $\langle
L_{X}/L_{\rm edd}\rangle$ should increase with black hole mass, though
only weakly for $M_{\bullet} \lesssim 3\times 10^{8}M_{\odot}$.  This
result is in tension with the observed down-sizing trend that the
average Eddington ratio is seen to decrease (albeit also weakly)
with $M_{\bullet}$.  However, one must keep in mind the enormous uncertainty in calculating the
luminosity of the accretion flow close to the black hole in a single
waveband based on the feeding rate on larger scales.  If disk outflows carry away most of the infalling mass before it reaches X-ray producing radii (e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}), then the
angular momentum of the infalling gas might also influence the X-ray luminosity, in a way
that could depend systematically on the stellar population and hence $M_{\bullet}$.

Perhaps a more readily addressable test is whether a steady-state accretion
picture can account for the the large scatter, typically of $2-3$ orders of magnitude, in $L_{X}/L_{\rm
  edd}$ at fixed $M_{\bullet}$.  Scatter could result from at 
least three effects: (1) instric scatter in the scaling relationships ($\Mbh-\sigma$ and $\Mbh-L_{\rm
  bulge}$) used to measure $\Mbh$; (2) the strong sensitivity of the accretion rate to
the stellar wind velocity, $\dot{M}/\dot{M}_{\rm edd} \propto
v_{w}^{-3.8(-2.4)}$ for core(cusp) galaxies, respectively
(Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]).  When 
combined with the significant dependence of $v_w$ on stochastic
intermittency in the star formation history (Appendix \ref{fig:NickPlot2}), this can lead to order of magnitude differences in $\dot{M}$; (3) possible differences angular momentum of the infalling gas and consequentiallly the
efficacy of disk outflows may vary from galaxy to galaxy at fixed
$\Mbh$.  Also potentially contributing is the order of magnitude difference, at fixed $v_w$ and $M_{\bullet}$, between $\dot{M}$ for core and cusp galaxies at fixed $v_w$ (Fig.~\ref{fig:bh_growth}) which could also lead to
scatter of $\dot{M}$ due to the variation in the Nuker stellar density
profile $\Gamma$ at fixed $M_{\bullet}$.  Any such differences in $\dot{M}$ are exacerbated by the expectation
that $L_{\rm X} \propto \dot{M}^{2} $ for radiatively inefficient
flows.

%Many of the upper
%limits on $L_{\rm X}$ at low $M_{\bullet}$ may in fact correspond to a
%separate ``thermally stable" population.  This possible dichotomy
%calls into question the use of simple extrapolations of power-law fits
%to the $L_{\rm X}(M_{\bullet}$) relationship to low $M_{\bullet}$ to
%constrain the occupation fraction of black holes in low mass galaxies
%(\citealt{Miller+15}).

@@
